Self Regulation Ontology Retest Analyses

### 2018-04-29 code/SRO_Retest_Analyses.Rmd
changing systematic diff analyses

@@

SEM CI calculation

```{r}
boot_df
```

>>>>

### Anova time effects

@@

### SEM CI calculation

Step 1: T = Grand mean + ICC * (Subject score - Grand mean)
Step 2: SEP = SD of both measurements * sqrt(1-ICC^2)

Here I calculate the proportion of subjects that move more than one standard error of prediction (SEP) in t2 compared to t1 for each measure.

This is odd to think about because the larger the ICC of a measure the smaller the SEP. So very small differences between the two time points can be categorized as 'meaningful' based on the tiny SEP.

What you are interested in is not necessarily whether individuals change at all between the two time points though. You want to know if this change is systematic in one direction.

Here I calculate the proportion of people showing 'meaningful' changes in one direction or the other. To integrate both of these direction I subtracted one from the other and filtered the variables that have more than 5% of the participants showing meaningful change in one direction over the other (so if a variable has 10 participants showing difference in one and 10 in the other direction this would cancel out but a variables with 15 people showing a positive and 5 negative change would remain).

```{r}
get_ind_ci = function(dv_var){
  matched = match_t1_t2(dv_var)
  grand_mean = mean(matched$score)
  grand_sd = sd(matched$score)
  dv_icc = get_icc(dv_var)
  sep = grand_sd * sqrt(1-(dv_icc^2))
  matched = matched %>% 
    spread(time, score) %>%
    rename("t1"="1", "t2"="2") %>%
    mutate(true_score = grand_mean+(dv_icc*(t1-grand_mean)),
           ci_up = true_score+(1.96*sep),
           ci_low = true_score-(1.96*sep),
           t2_above_ci = ifelse(t2>ci_up,1,0),
           t2_below_ci = ifelse(t2<ci_low,1, 0))
    return(matched)
}

get_prop_out_ci = function(dv_var){
  get_ind_ci(dv_var) %>%
    summarise(prop_above_ci = sum(t2_above_ci)/n(),
              prop_below_ci = sum(t2_below_ci/n()))
}

#Create df of point estimate reliabilities
ind_ci_df <- data.frame(prop_above_ci = rep(NA, length(numeric_cols)),
                        prop_below_ci = rep(NA, length(numeric_cols)))

row.names(ind_ci_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
    ind_ci_df[numeric_cols[i], 'prop_above_ci'] <- get_prop_out_ci(numeric_cols[i])$prop_above_ci 
    ind_ci_df[numeric_cols[i], 'prop_below_ci'] <- get_prop_out_ci(numeric_cols[i])$prop_below_ci 
}

ind_ci_df$dv = row.names(ind_ci_df)
row.names(ind_ci_df) = seq(1:nrow(ind_ci_df))
ind_ci_df$task = 'task'
ind_ci_df[grep('survey', ind_ci_df$dv), 'task'] = 'survey'
ind_ci_df[grep('holt', ind_ci_df$dv), 'task'] = "task"
ind_ci_df = ind_ci_df %>%
  select(dv, task, prop_above_ci, prop_below_ci)

mean_diff_df = ind_ci_df %>%
  mutate(prop_one_direction = prop_above_ci-prop_below_ci) %>%
  filter(abs(prop_one_direction)>0.05) %>%
  arrange(-abs(prop_one_direction))

mean_diff_df

write.csv(mean_diff_df, '/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/tables/mean_diff_df.csv')
```

This doesn't make it easier to interpret whether it is a performance improvement or decline. In fact 'performance' isn't even necessarily the correct term here. Using the valence assignments of the measures look at whether it translates to an increase or decrease in self-regulation. Of the 66 this makes sense for 54 variables.

```{r warning=FALSE, message=FALSE}
mean_diff_df = read.csv('/Users/zeynepenkavi/Dropbox/PoldrackLab/SRO_Retest_Analyses/output/tables/mean_diff_df.csv')

table(as.character(mean_diff_df$valence), useNA = "ifany")

mean_diff_df %>%
  mutate(sc_increase = ifelse(prop_one_direction>0&valence=="Pos",1,ifelse(prop_one_direction<0&valence=="Neg",1,0)),
         sc_decrease = ifelse(prop_one_direction>0&valence=="Neg",1,ifelse(prop_one_direction<0&valence=="Pos",1,0))) %>%
summarise(sum_sc_increase = sum(sc_increase,na.rm=T),
          sum_sc_decrease = sum(sc_decrease,na.rm=T))

Variables that show more than 5% difference in a direction do not depend on whether they translate to a self control increase or decrease.

```{r}
chisq.test(matrix(data = c(23,54-23,31,54-31), nrow=2))
```

Do they differ depending on whether they are task or survey variables? No.

```{r}
chisq.test(matrix(data = c(5,74-5,61,372-61), nrow=2))
```

```{r}
rm(mean_diff_df)
```
